---
title: "Impact of Research on Univerisity Rankings"
author: "Thomas Billington, Steven Bottone, Nart Hakij, and Joe Manley"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction:
* Many schools either brand themselves as teaching, professional, or research focused universities.  
* By measuring the trends of institutions across the globe we can determine if their ability to conduct meaningful research correlate to its statute.  
* Goal: Does a university’s research efforts have an effect on its global ranking?  
* Response variable: Global ranking  
* Explanatory variable: Research score (out of 100)  
* Observations: 1695 universities  

## Analysis: 
To better understand the relationship, a thorough statistical analysis was conducted.    
```{r chunk1, echo=FALSE, warning=FALSE, comment=NA, message = FALSE}
library(ggplot2)
data = read.csv("rankings.csv")
data = na.omit(data)
data$University.Rank <- as.numeric(data$University.Rank)
data$Research.Score <- as.numeric(data$Research.Score)

model <- lm(University.Rank ~ Research.Score, data = data)

# https://www.kaggle.com/datasets/alitaqi000/world-university-rankings-2023
plot = ggplot(data, aes(x = Research.Score, y = University.Rank)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE, color = "red") + ylim(0, NA) +
  labs(y = "University Rank", x = "Research Score") +
  ggtitle("University Rank vs Research Score")

plot


correlation = cor(data$University.Rank, data$Research.Score)
# correlation
```
From the graph we can assume a strong negative linear relationship between research score and university ranking. We see that as research score increases, rank decreases(goes towards better rank).  
  
More information can be observed from better understanding the linear model:
```{r chunk2, echo=FALSE, warning=FALSE, comment=NA, message = FALSE}
coefficients <- coef(model)
intercept <- coefficients[1]
slope <- coefficients[2]

summary(model)
```
From this summary output we can note a few key details about the model.

* Ŷ = 1346.679 + (-21.23218)*X
* Ŷ is university ranking
* X is research score of university  

## Statistical Tests and Inference:  
  
Hypothesis Testing:  

* H0: There is no association between university ranking and research score  
* H1: There is an association between university ranking and research score    
  
The p-value was 2.2e^-16 which is much lower than the significance level of 0.05. Therefore, we can reject the null hypothesis, indicating that there is a strong statistical evidence that when university ranking decreases or gets closer to rank 1, research score increases.  
  
Correlation:    

* The R^2 value indicates that 54.2% of the variation in university rank can be explained by the research score.  
* The estimated residual standard error is s = 331.8 on 1695 degrees of freedom. This indicates that there is some variability in the models predictions.    
* Correlation coefficient between University.Rank and Research.Score: -0.736. This indicates a strong negative correlation, as one variable increases the other decreases.  






Confidence Intervals:
```{r}
confidence_interval <- confint(model, level = 0.95)
confidence_interval
```
Prediction stuff
```{r}
new_data <- data.frame(Research.Score = 60)
pred_int <- predict(model, newdata = new_data, interval = "prediction", level = 0.95)
pred_int
lower_bounds <- pred_int[, 1]
upper_bounds <- pred_int[, 2]

# Set negative lower bounds to 0
lower_bounds <- pmax(lower_bounds, 0)

# Set negative upper bounds to 0
upper_bounds <- pmax(upper_bounds, 0)

# Combine the modified lower and upper bounds into a matrix
pred_int <- cbind(lower_bounds, upper_bounds)
pred_int
```

